---
title: "ESM232_FinalProject"
author: "Emma Mendonsa & Claire Powers"
date: "April 29, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Overarching Question: 
- How will climate change impacts on bark beetle populations and tree populations influence forest carbon sequestration?

- Approach: This larger model relies on three submodels: 1) bark beetle population, 2) tree population, and 3) tree volume burned. 

- Expectations: We expect temperature changes to influence the success of both populations. Water availability has been show to be influential in tree defense from bark beetles. Carbon sequestration will incorporate the volume of dead trees and forest NPP to provide information about how much carbon is sequestered in the forest following a fire. 

- Extra component: The probility of fire and the fire's severity will be a user controlled input. Three fire severity levels will be pre-defined and selected by the user (high, medium, and low).

###Rationale for Model Importance:
INSERT 1-2 paragraphs here...


##Beginning Model Inputs: 
```{r}



```

