---
title: "Sensitivity_Analysis_pines"
author: "Claire Powers & Emma Mendonsa"
date: "June 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE, message=FALSE, error=FALSE}
library(tidyverse)
library(tibble)
library(reshape2)
library(sensitivity)
library(pse)
library(popbio)


source("../R/SubModel2_PineModel.R")
```

###Goals:
1. Perform a sensitivity analysis on the growth of whitebark pines
  -->> If all else is held constant, is the variability in the pine population
2. Perform a sensitivity analysis on the survival and fecundity of whitebark pines
  -->> We want to see how survival changes with intervention (fire, beetles, etc), but we are also uncertain about the fecundity of the whitebark pines.


####Sensitivity Analysis: Growth
Data Setup: 
```{r}
nsample=200
stages = c("<1cm","1-10cm","10-20cm","20-40cm",">40cm")
initial_pop = c(100,100,50,50,50)
time = 2

# Pine Fecundity
F0 = 0 
F1 = 0 
F2 = 0.0742
F3 = 0.2536 
F4 = 1.6902

# Pine Survival
p0 = 0.9353
p1 = 0.9827
p2 = 0.9720
p3 = 0.9806
p4 = 0.9959

# Pine Growth
g01 = 0.0018
g12 = 0.0104
g23 = 0.0144
g34 = 0.0144

# Create our two samples for Sobel
# Growth
gs1 = cbind.data.frame(g01 = runif(min = 0.0010, max=0.0025, n=nsample), # Published g01 value = 0.0018
                       g12 = runif(min = 0.0005, max=0.0015, n=nsample),  # Published g12 value = 0.0104
                       g23 = runif(min = 0.0100, max=0.0200, n=nsample),  # Published g23 value = 0.0144
                       g34 = runif(min = 0.0100, max=0.0200, n=nsample))  # Published g34 value = 0.0144

gs2 = cbind.data.frame(g01 = runif(min = 0.0010, max=0.0025, n=nsample), 
                       g12 = runif(min = 0.0005, max=0.0015, n=nsample),  
                       g23 = runif(min = 0.0100, max=0.0200, n=nsample),  
                       g34 = runif(min = 0.0100, max=0.0200, n=nsample))

# now include uncertainty in our Fertility
fs1 = cbind.data.frame(F0=rep(F0, times = nsample), 
                       F1=rep(F1, times = nsample),
                       F2=rep(F2, times = nsample),
                       F3=rep(F3, times = nsample),
                       F4=rep(F4, times = nsample))

fs2 = cbind.data.frame(F0=rep(F0, times = nsample), 
                       F1=rep(F1, times = nsample),
                       F2=rep(F2, times = nsample),
                       F3=rep(F3, times = nsample),
                       F4=rep(F4, times = nsample))

# now include uncertainty in our Survivability
ps1 = cbind.data.frame(p0=rep(p0, times = nsample), 
                       p1=rep(p1, times = nsample),
                       p2=rep(p2, times = nsample),
                       p3=rep(p3, times = nsample),
                       p4=rep(p4, times = nsample))

ps2 = cbind.data.frame(p0=rep(p0, times = nsample), 
                       p1=rep(p1, times = nsample),
                       p2=rep(p2, times = nsample),
                       p3=rep(p3, times = nsample),
                       p4=rep(p4, times = nsample))


# put survivability and fertility together
all_growth1 = cbind.data.frame(gs1,fs1,ps1)
all_growth2 = cbind.data.frame(gs2,fs2,ps2)
```


Sobel Run:
```{r}
# get sobel samples
sens_growth = sobol2007(model = NULL, all_growth1, all_growth2, nboot = 300)

head(sens_growth$X)
nsim=nrow(sens_growth$X)

# Run over 91 years to see final population
# use as numeric to extract parameters
res = rep(0, times=nsim)
for (i in 1:nsim) {
  tmp = whitePine_matrix_model(stages = stages, 
                               fertility = as.numeric(sens_growth$X[i,c("F0","F1","F2","F3", "F4")]), 
                               survival = as.numeric(sens_growth$X[i,c("p0","p1","p2", "p3", "p4")]), 
                               growth = as.numeric(sens_growth$X[i,c("g01", "g12","g23","g34")]), 
                               initial_pop = initial_pop, 
                               time = 91)
  res[i] = tmp$total_pop[91]
  print(i)
}

# give our results to sensitivity structure

sens_growth = sensitivity::tell(sens_growth, res)

# look at results. 

sens_growth$S  ## The results show that the growth in the 1st age class is the most important.
sens_growth$T

tmp = cbind.data.frame(sens_micro$X, pop_total=sens_micro$y)

ggplot(tmp, aes(x = g01, pop_total))+geom_point()
ggplot(tmp, aes(x = g12, pop_total))+geom_point()
ggplot(tmp, aes(x = g23, pop_total))+geom_point()
ggplot(tmp, aes(x = g34, pop_total))+geom_point()
```


####Sensitivity Analysis: Fecundity 
Data Setup: 
```{r}
nsample=200
stages = c("<1cm","1-10cm","10-20cm","20-40cm",">40cm")
initial_pop = c(100,100,50,50,50)
time = 2

# Pine Fecundity
F0 = 0 
F1 = 0 
F2 = 0.0742
F3 = 0.2536 
F4 = 1.6902

# Pine Survival
p0 = 0.9353
p1 = 0.9827
p2 = 0.9720
p3 = 0.9806
p4 = 0.9959

# Pine Growth
g01 = 0.0018
g12 = 0.0104
g23 = 0.0144
g34 = 0.0144

# Create our two samples for Sobel
# Growth

# now include uncertainty in our Fertility
fs1 = cbind.data.frame(F0=rep(F0, times = nsample), 
                       F1=runif(min = 0.0010, max=0.0025, n=nsample), # 0
                       F2=runif(min = 0, max=0.5, n=nsample), # 0.0742
                       F3=runif(min = 0, max=1, n=nsample), # 0.2536
                       F4=runif(min = 1, max=2, n=nsample)) # 1.6902

fs2 = cbind.data.frame(F0=rep(F0, times = nsample), 
                       F1=runif(min = 0.0010, max=0.0025, n=nsample), # 0
                       F2=runif(min = 0, max=0.5, n=nsample), # 0.0742
                       F3=runif(min = 0, max=1, n=nsample), # 0.2536
                       F4=runif(min = 1, max=2, n=nsample))

# now include uncertainty in our Survivability
gs1 = cbind.data.frame(g01=rep(g01, times = nsample), 
                       g12=rep(g12, times = nsample),
                       g23=rep(g23, times = nsample),
                       g34=rep(g34, times = nsample))

gs2 = cbind.data.frame(g01=rep(g01, times = nsample), 
                       g12=rep(g12, times = nsample),
                       g23=rep(g23, times = nsample),
                       g34=rep(g34, times = nsample))

# Survival
ps1 = cbind.data.frame(p0=rep(p0, times = nsample), 
                       p1=rep(p1, times = nsample),
                       p2=rep(p2, times = nsample),
                       p3=rep(p3, times = nsample),
                       p4=rep(p4, times = nsample))

ps2 = cbind.data.frame(p0=rep(p0, times = nsample), 
                       p1=rep(p1, times = nsample),
                       p2=rep(p2, times = nsample),
                       p3=rep(p3, times = nsample),
                       p4=rep(p4, times = nsample))


# put survivability and fertility together
all_fecundity1 = cbind.data.frame(gs1,fs1,ps1)
all_fecundity2 = cbind.data.frame(gs2,fs2,ps2)
```


Sobel Run:
```{r}
# get sobel samples
sens_fecundity = sobol2007(model = NULL, all_fecundity1, all_fecundity2, nboot = 300)

head(sens_fecundity$X)
nsim=nrow(sens_fecundity$X)

# Run over 91 years to see final population
# use as numeric to extract parameters
res = rep(0, times=nsim)
for (i in 1:nsim) {
  tmp = whitePine_matrix_model(stages = stages, 
                               fertility = as.numeric(sens_fecundity$X[i,c("F0","F1","F2","F3", "F4")]), 
                               survival = as.numeric(sens_fecundity$X[i,c("p0","p1","p2", "p3", "p4")]), 
                               growth = as.numeric(sens_fecundity$X[i,c("g01", "g12","g23","g34")]), 
                               initial_pop = initial_pop, 
                               time = 91)
  res[i] = tmp$total_pop[91]
  print(i)
}

# give our results to sensitivity structure
sens_fec_results = sensitivity::tell(sens_fecundity, res)

# look at results. 

sens_fec_results$S  ## The results show that the fecundity in the 4th age class (F3) is the most important.
sens_fec_results$T

sens_fec_df = cbind.data.frame(sens_fec_results$X, pop_total=sens_fec_results$y)

ggplot(sens_fec_df, aes(x = F1, pop_total))+geom_point()
ggplot(sens_fec_df, aes(x = F2, pop_total))+geom_point()
ggplot(sens_fec_df, aes(x = F3, pop_total))+geom_point()
ggplot(sens_fec_df, aes(x = F4, pop_total))+geom_point()
```


####Sensitivity Analysis: Survival 
Data Setup: 
```{r}

nsample=200
stages = c("<1cm","1-10cm","10-20cm","20-40cm",">40cm")
initial_pop = c(100,100,50,50,50)
time = 2

# Pine Fecundity
F0 = 0 
F1 = 0 
F2 = 0.0742
F3 = 0.2536 
F4 = 1.6902

# Pine Survival
p0 = 0.9353
p1 = 0.9827
p2 = 0.9720
p3 = 0.9806
p4 = 0.9959

# Pine Growth
g01 = 0.0018
g12 = 0.0104
g23 = 0.0144
g34 = 0.0144

# Create our two samples for Sobel
# Growth
ps1 = cbind.data.frame(p0 = rnorm(mean=0.9353, sd=0.15, n=nsample), 
                       p1 = rnorm(mean=0.9827, sd=0.15, n=nsample),
                       p2 = rnorm(mean = 0.9720, sd=0.15, n=nsample),
                       p3 = rnorm(mean = 0.9806, sd=0.15, n=nsample),                       
                       p4 = rnorm(mean = 0.9959, sd=0.15, n=nsample))

p2 = cbind.data.frame(p0 = rnorm(mean=0.9353, sd=0.15, n=nsample), 
                       p1 = rnorm(mean=0.9827, sd=0.15, n=nsample),
                       p2 = rnorm(mean = 0.9720, sd=0.15, n=nsample),
                       p3 = rnorm(mean = 0.9806, sd=0.15, n=nsample),                       
                       p4 = rnorm(mean = 0.9959, sd=0.15, n=nsample))

# now include uncertainty in our Fertility
fs1 = cbind.data.frame(F0=rep(F0, times = nsample), 
                       F1=rep(F1, times = nsample),
                       F2=rep(F2, times = nsample),
                       F3=rep(F3, times = nsample),
                       F4=rep(F4, times = nsample))

fs2 = cbind.data.frame(F0=rep(F0, times = nsample), 
                       F1=rep(F1, times = nsample),
                       F2=rep(F2, times = nsample),
                       F3=rep(F3, times = nsample),
                       F4=rep(F4, times = nsample))

# now include uncertainty in our Survivability
gs1 = cbind.data.frame(g01=rep(g01, times = nsample), 
                       g12=rep(g12, times = nsample),
                       g23=rep(g23, times = nsample),
                       g34=rep(g34, times = nsample))

gs2 = cbind.data.frame(g01=rep(g01, times = nsample), 
                       g12=rep(g12, times = nsample),
                       g23=rep(g23, times = nsample),
                       g34=rep(g34, times = nsample))


# put survivability and fertility together
all_survival1 = cbind.data.frame(gs1,fs1,ps1)
all_survival2 = cbind.data.frame(gs2,fs2,ps2)
```


Sobel Run:
```{r}
# get sobel samples
sens_survival = sobol2007(model = NULL, all_survival1, all_survival2, nboot = 300)

head(sens_survival$X)
nsim=nrow(sens_survival$X)

# Run over 20 years to see final population
# use as numeric to extract parameters
res = rep(0, times = nsim)
for (i in 1:nsim) {
  tmp = whitePine_matrix_model(stages = stages, 
                               fertility = as.numeric(sens_survival$X[i,c("F0","F1","F2","F3", "F4")]), 
                               survival = as.numeric(sens_survival$X[i,c("p0","p1","p2", "p3", "p4")]), 
                               growth = as.numeric(sens_survival$X[i,c("g01", "g12","g23","g34")]), 
                               initial_pop = initial_pop, 
                               time = 91)
  res[i] = tmp$total_pop[91]
  print(i)
}

# give our results to sensitivity structure
sens_survival_results = sensitivity::tell(sens_survival, res)

# look at results. 

sens_survival_results$S  ## The results show that 
sens_survival_results$T

sens_survival_df = cbind.data.frame(sens_survival_results$X, pop_total=sens_survival_results$y)

ggplot(sens_survival_df, aes(x = p0, pop_total))+geom_point()
ggplot(sens_survival_df, aes(x = p1, pop_total))+geom_point()
ggplot(sens_survival_df, aes(x = p2, pop_total))+geom_point()
ggplot(sens_survival_df, aes(x = p3, pop_total))+geom_point()
ggplot(sens_survival_df, aes(x = p4, pop_total))+geom_point()
```

